---
title: "Anova project--"
author: "Mo'men Mohamed"
date: "March 31, 2019"
output: html_document
---



#This Mark down for anova 
### ---> for all types of anova 

loading the libraries 

```{r lib,message=F,warning=FALSE}
library(tidyverse)
library(knitr)
library(pwr)
library(car)
library(sjstats)
library(psych)
library(ggplot2)
```

loading the data sets --> All data sets 

```{r data}
bs<-read.csv("./data sets/bn 1 anova.csv")

names(bs)<-c("group","obs")

head(bs)%>%kable("markdown") # it is a longtudinal data set 


```

First of all lets see the desired sample size 

```{r sample size}

#intial anova with our sample 

aov(formula = obs~group,data = bs)%>%summary
betvar<-25.24/2 #mean sq
withvar<-19.67/8 #mean sq
pwr.anova.test(k = 3,f = .25,sig.level = .05,power = .8)

#here we see that we need arround 53 person in each group to acheive the power of .8

#but we pretend that we have this size !!

```


Now lets look at the acurracy , missing and outlier 

```{r screen}
summary(bs)

#no missing 

#see if there are any outlier 

max<-qnorm(p = c(.99),mean = mean(bs$obs),sd = sd(bs$obs))
min<-qnorm(p = c(.01),mean = mean(bs$obs),sd = sd(bs$obs))

bs[which(bs$obs>max|bs$obs<min),]

# there is no outliers !!

```


assumption of normality 


```{r norm}
#running fake regression 
rdata<-rnorm(n = nrow(bs))
fake<-lm(rdata~bs$obs)
plot(fake)

qqPlot(fake)
```

the data is linear and normal but it is not enough to judge on homoscdatisity and homogeniety 

```{r homogeniety}
leveneTest(bs$obs~bs$group)

```

great ,, the data is homogeneic

Now lets run anova 


```{r anova}
mod<-bs%>%with(aov(obs~group))

eta_sq(mod) # we have eta square = 0.56


```

as the f statistic is significant , then we have to do post hoc test to know where is exactly the difference , and this pair wise is corrected using be

```{r post}
t<-bs%>%with(
        pairwise.t.test(x = obs,group,p.adjust.method = "bonferroni",paired = F)
)
t


bs%>%group_by(group)%>%summarize(mean=mean(obs))%>%kable("markdown")
```

lets graphing it

```{r graph}

errordata<-bs%>%group_by(group)%>%summarize(mean=mean(obs),max=qnorm(.95,mean = mean(obs),sd=sd(obs)),min=qnorm(.05,mean = mean(obs),sd=sd(obs)))

errordata%>%ggplot(aes(ymin=min,ymax=max,x=group))+geom_pointrange(aes(x = group,y = mean,color=group))+geom_errorbar()
        
```
