---
title: "AZZA"
author: "Mo'men Mohamed"
date: "February 8, 2019"
output: html_document
---

**libraries**

```{r libraries,message=FALSE}
library(readxl)

library(magrittr)

library(MVN)

library(tidyr)

library(dplyr)

library(psych)

library(ggplot2)

library(car)
```

#Test Validation And Reliability

##Adjustment test

**importing the data**

```{r import}
adjust_score<-read_excel("./data sets/azza/azza-adjust.xlsx",sheet = 1)
adjust_judge<-read_excel("./data sets/azza/azza-adjust.xlsx",sheet = 2)
```

**Knowing the class**

```{r class}
apply(adjust_score,2,class)

class(adjust_score)

adjust_mat<-adjust_score[,-1]
```

**decribing the data **

```{r decribe }
round(apply(adjust_mat,2,function(x){
        c(max=max(x),
          min=min(x),
          na=sum(is.na(x)),
          mean=mean(x),
          median=median(x),
          sd=sd(x))
}),2)


```

**normality**

```{r normality,error=TRUE}

#qqplot to residuals 

resid<-lm(A1~.,adjust_mat)$resid
resid<-data.frame(resid=resid)
resid%>%ggplot(aes(x = resid))+geom_histogram()
qqPlot(x = resid$resid)

resid2<-lm(A1~.,adjust_mat[-c(30,54),])$resid
resid2<-data.frame(resid=resid2)
resid2%>%ggplot(aes(x = resid))+geom_histogram()
qqPlot(x = resid2$resid)

#shapiro test

apply(adjust_mat,2,shapiro.test)

apply(adjust_mat,2,function(x){
        tryCatch(shapiro.test(x = x),error=function(err){NA})
})

#Mardia test

mardia(adjust_mat)


```

**Unfortunatly the data is so singular **

**testing the additivity **

```{r additivity}
cor(adjust_mat)
```


**Ok , I decided to analyse the SME data to see the agreement on the Items**

```{r see the data}
head(adjust_judge)
```

**calculating the percent of agreement for each item **

```{r percent calculation}
adjust_judge$agree<-apply(adjust_judge[,-1],1,function(x){
        sum(x)/length(x)
})

adjust_judge
```

**assigning items lower than *.8***

```{r lower .8}
adjust_judge[adjust_judge$agree<.8,]
```

**calculating binomial test **

```{r binomial test}
adjust_judge$P.Value<-rep(NA,nrow(adjust_judge))
for(i in 1:nrow(adjust_judge)){
        round(adjust_judge$P.Value[i]<-binom.test(x = rowSums(adjust_judge[i,c(2:11)]),n = 10,p = .8,alternative = "less")$p.value,2)
}

adjust_judge[,c(1,13)]

```