---
title: "Between subject MANOVA Tutorial"
author: "Mo'men Mohamed"
date: "April 5, 2019"
output: html_document
---

This a tutorial for doing between subject MANOVA

you are welcome to contact me for any advice or recommendation at: vet.m.mohamed@gmail.com

for Complete tutorial and data set see [here](http://statstools.com/learn/advanced-statistics/)


first of all lets load libraries which we are going to use in this Analysis and clearing the work space

```{r,message=F,warning=F}
library(tidyverse)
library(car)
library(multcomp)
library(ez)
library(MOTE)
library(psych)
```

Then lets import the data set

```{r}
data<-read.csv("./data sets/9 manova.csv")
head(data)
```

lets take the analysis variables out 

```{r}
data<-data[,c("FEM","MASC","ESTEEM","ATTROLE","NEUROTIC")]
```

adding individual index

```{r}
data<-data%>%mutate(idx=1:n())
```

Now lets test the assumptions and the appropriateness of the data to MANOVA


We will start by accuracy and missing data

```{r}
summary(data)
```

In Neurotic we have a large numbers of Na's 

So we have to see the percentage of the missing data 

```{r}
apply(data,2,function(x){
        sum(is.na(x))/length(x)
})
```
We have here 17.8% of data in neurotic are missing so we can't do imputation 


Now lets test for outliers using maharanis test

```{r}
mah<-mahalanobis(x = data[,-c(1,2,6)],
                 center = colMeans(data[,-c(1,2,6)],na.rm = T),cov = cov(data[,-c(1,2,6)],use = "pairwise.complete.obs"))
```

determining the cutoff point through chi square distribution 

```{r}
cut<-qchisq(p = .99,df = 2)

```

summary of the outliers

```{r}
summary(mah<cut)
```

we have 9 outliers and we have to remove them 

```{r}
data<-data[which(mah<cut),]
```

Now lets move to the second phase :

testing the normality - linearity - Homogeneity - Homoscdasticity

```{r}
par(mfrow=c(2,2))
qqnorm(data$ESTEEM%>%jitter(factor = 5),main = "QQplot for self esteem")
qqnorm(data$ATTROLE%>%jitter(factor = 5),main = "QQplot for Attitude to roles")
qqnorm(data$NEUROTIC%>%jitter(factor = 5),main = "QQplot for neurotic")

```

Other way : making fake data for fake regression to test the assumption in multivariate way

```{r}
set.seed(345)
rfake<-rchisq(n = nrow(data),7)
```
the df is arbitrary number 

```{r}
fakelm<-data%>%with(lm(rfake~ESTEEM+ATTROLE+NEUROTIC))
```

Now lets Extract the residual and fitted values 

```{r}
stdresid<-resid(fakelm)%>%scale()
stdfit<-fitted(fakelm)%>%scale()
```

start with normality 
we can check it using qqnorm plot

```{r}
qqPlot(stdresid)
```

fair enough !! got good data here 

Now its time for homogeneity and homoscdasticity

```{r}
plot(stdfit,stdresid)
abline(h = 0)
abline(v = 0)
```

good , the data here is some king of homogeneity 

 but we still need use test like levene to prove it 

```{r}
data%>%with(leveneTest(ESTEEM~FEM*MASC)) #testing for self Esteem
data%>%with(leveneTest(ATTROLE~FEM*MASC)) #testing for Attitude toword female roles
data%>%with(leveneTest(NEUROTIC~FEM*MASC)) # testing for neuroticism
```

the size of the sample Affect greatly on the significance level 
so we will choose p.value =.001

all Homogeneity tests is p>.001

so we will assume the homogeneity here 

now lets take more steps further toward manova analysis 

```{r}
DV = cbind(data$ESTEEM, data$ATTROLE, data$NEUROTIC)#combining the dependent variables together

man<-data%>%with(lm(DV~FEM*MASC,data = data,
                    contrasts = list(FEM=contr.sum,MASC=contr.sum)))

manout<-Manova(mod = man,"III") #type three SS

summary(manout,multivariate = T)
```

here we see the the significant main effect of the IVs 

the effect size here = 1- wilk's lambda 

```{r}
1-0.917827 #effect size of FEM
1-0.8237555 #effect size of MASC
1-0.9933737 #effect size of interaction
```


here we will do post-hoc test --> single anova for each DV

```{r,warning=F,message=F}
#self esteem
data%>%with(ezANOVA(data = data,dv = ESTEEM,
                    wid = idx,between = c(FEM,MASC),type = 3))

#Attitude
data%>%with(ezANOVA(data = data,dv = ATTROLE,
                    wid = idx,between = c(FEM,MASC),type = 3))

#Neurotic

data%>%with(ezANOVA(data = na.omit(data),dv = NEUROTIC,
                    wid = idx,between = c(FEM,MASC),type = 3))
```

thankful the IVs is two level only and there is no interaction effect (except in self esteem)
so the p-value in anova is enough and Explain only the mean 

but we still need to calculate the effect size of each sig. different means using MOTE library or Cohen_d() function in psych package

using Cohen.d()

```{r}
data[,-c(2,6)]%>%cohen.d(group = "FEM")

data[,-c(1,6)]%>%cohen.d(group = "MASC")
```

using d.ind.t()

```{r}
data%>%group_by(FEM)%>%summarize(mean=mean(ATTROLE,na.rm = T),
                                      sd=sd(ATTROLE,na.rm = T),num=n())

d.ind.t(m1 = 36.0,m2 = 32.9,sd1 = 6.80,sd2 = 6.41,n1 = 202,n2 = 90)$d
```

we here see that the effect size using Cohen.d() is so close to d.ind.t() for attribute on each group of FEM
but Cohen.d() here is more easier 


HOPE TO FIND IT USEFULL 

THANKS FOR BEING HERE

REGARDS


