---
title: "BS Two way anova"
author: "Mo'men Mohamed"
date: "April 1, 2019"
output: html_document
---

loading the libraries 

```{r lib load}
library(tidyverse)
library(psych)
library(ez)
library(knitr)
library(reshape)
library(car)

```

loading the data 

```{r data}
bs2<-read.csv("./data sets/bn 2 anova.csv")
bs2<-bs2%>%mutate(year=factor(year),type=factor(type))
head(bs2)%>%kable("markdown")

#know the factors levels 

apply(bs2[,1:2],2,unique)
```

we have two factors :
years : two years 
sport : 5 levels 

Note: testing power and sample size using G-power App
and for small effect size , interaction df (2-1)*(5-1)=4
and nuber of group (conditions) = 10 , with power .8 --> we need 126 sample size 

Lets see the data summary 

```{r summary}
bs2%>%summary()


# there is no out of bound cases and no missing value 

```

lets test outleir and normality 

```{r out-norm}
prp<-(pnorm(bs2$money,mean = mean(bs2$money,na.rm = T),sd = sd(bs2$money,na.rm = T)))

bs2<-bs2[which(prp<.99&prp>.01),] #Exclude outliers

#normality

faker<-rchisq(n = nrow(bs2),df = 7)

fakelm<-lm(faker~.,data = bs2)

stdresid<-rstudent(fakelm)
stdfit<-scale(fitted(fakelm))

hist(stdresid)

qqPlot(stdresid) # we have here a quadratic functuion 
# but we still do the analysis 

plot(stdresid,stdfit)%>%abline(h = 0)%>%abline(v = 0)

#we have a serious problem in homogeneity and homoscadisticity 
```

testing the homogeneity

```{r homo}
bs2%>%na.omit()%>%
        with(leveneTest(money~type*year)) #Not homogeneic data
```

Running anova

```{r anova}
bs2<-bs2%>%mutate(idx=1:nrow(bs2))
head(bs2)


bs2%>%with(ezANOVA(data = bs2,dv =money,wid = idx,between = c(year,type),type = 1))
```

the interaction is not significant 

So, we will test the main effect only 

between years and between type of game 


```{r post hoc}
bs2%>%with(pairwise.t.test(x = money,g = type,p.adjust.method = "bonferroni",paired = F,equal.var=F))

bs2%>%with(pairwise.t.test(x = money,g = year,p.adjust.method = "bonferroni",paired = F,equal.var=F))

```

