---
title: "Poly IRT"
author: "Mo'men Mohamed"
date: "March 30, 2019"
output: html_document
---

```{r lib}
library(mirt)
library(tidyverse)
library(knitr)
library(ltm)
library(psych)
```

loading the data 

```{r load}
data("LSAT")
kable(LSAT%>%head,"markdown")

poly<-read.csv("./data sets/poly irt.csv")

poly<-na.omit(poly)
head(poly)%>%kable("markdown")
```
Working on LSAT dichotomus data 

loading the model

```{r dich model}
dimodel<-ltm(LSAT~z1,IRT.param = T)

# see the model summary 

summary(dimodel)

#looking at the coefficients sololy 

coef(dimodel)

#lets Now plot the model ICC

plot(dimodel,type="ICC")
plot(dimodel,type="ICC",1)
plot(dimodel,type="ICC",2)

#Now IIC plotting 

plot(dimodel,type="IIC",0)
plot(dimodel,type="IIC",1)
plot(dimodel,type="IIC",2)
plot(dimodel,type="IIC",3)
plot(dimodel,type="IIC",4)
plot(dimodel,type="IIC",5)


```


we see that most information is is arround -1 z and from the curve we can see the information and goodness of each paramater to each item 

Now lets look to the model in deep 

looking to the factor scores , person fit , item fit 

```{r deep}

##seeing the pattern of responses and z value of the respondant theta 

ltm::factor.scores(dimodel)

##See each person pattern and his theta level and propapility 

person.fit(dimodel)

## Now see the propapility of the item doesn't fit the model --- > we want it non significant 

item.fit(dimodel)

```


Now lets look at the 3 parameter model and see it it makes different !!


```{r 3p}
trimodel<-tpm(data = LSAT,IRT.param = T,type = "latent.trait")
#see the summary

summary(trimodel)

#see coef

coef(trimodel)

#plotting

plot(trimodel,type="ICC")

plot(trimodel,type="IIC",0)


#looking at the factor score

ltm::factor.scores(trimodel)

#looking at person score

person.fit(trimodel)

#looking at item fit 

item.fit(trimodel)
```

Now lets test the difference between the two model

```{r test}
anova(dimodel,trimodel)

```

3 paramater model doesn't add any information , so we will keep model 2

NOW its time to work on polytomus model 

we will treat each factor alone : we hypothsize that the item unidimentional 

```{r ploy}
#reverse item 9

poly$Q99_9<-8-poly$Q99_9

#Now separate each factor items 

factor1<-poly[,c(1, 4, 5, 6, 9)]
factor2<-poly[,c(2, 3, 7, 8, 10)]

#modeling factor 1
polymodel1<-mirt(factor1,model = 1, #for one factor 
                 itemtype= "gpcm",#generalized partial Credit model
                 technical = list(removeEmptyRows=TRUE)) 

# see the summary 

summary(polymodel1)

#seeing the coef
coef(polymodel1,IRTpars=T) #parameter of each choice 

#plotting the model

itemplot(polymodel1,1,type="trace")# ICC
itemplot(polymodel1,2,type="trace")# ICC
itemplot(polymodel1,3,type="trace")# ICC
itemplot(polymodel1,4,type="trace")# ICC
itemplot(polymodel1,5,type="trace")# ICC


itemplot(polymodel1,1,type="info") #IIC for each item
itemplot(polymodel1,2,type="info")
itemplot(polymodel1,3,type="info")
itemplot(polymodel1,4,type="info")
itemplot(polymodel1,5,type="info")


plot(polymodel1,type="trace") #plotting all items ICC
plot(polymodel1,type="info") #plotting all IIC



```    

Now lets see factor scores , person fit , item fit 

```{r deeb}
fscores(polymodel1)
personfit(polymodel1)
itemfit(polymodel1)
```

Now we will do the same for factor2


```{r factor2}
#modeling factor 2
polymodel2<-mirt(factor2,model = 1, #for one factor 
                 itemtype= "gpcm",#generalized partial Credit model
                 technical = list(removeEmptyRows=TRUE)) 

# see the summary 

summary(polymodel2)

#seeing the coef
coef(polymodel2,IRTpars=T) #parameter of each choice 

#plotting the model

itemplot(polymodel2,1,type="trace")# ICC
itemplot(polymodel2,2,type="trace")# ICC
itemplot(polymodel2,3,type="trace")# ICC
itemplot(polymodel2,4,type="trace")# ICC
itemplot(polymodel2,5,type="trace")# ICC


itemplot(polymodel2,1,type="info") #IIC for each item
itemplot(polymodel2,2,type="info")
itemplot(polymodel2,3,type="info")
itemplot(polymodel2,4,type="info")
itemplot(polymodel2,5,type="info")


plot(polymodel2,type="trace") #plotting all items ICC
plot(polymodel2,type="info") #plotting all IIC



fscores(polymodel2)
personfit(polymodel2)
itemfit(polymodel2)


```

