---
title: "EFA Tutorial"
author: "Mo'men Mohamed"
date: "March 5, 2019"
output: html_document
---


###This is An EFA Project explaining the practice and skills needed

###Notes:
####that the data used from [statstool](http://statstools.com/) and the tutorial guided by their tutorial 

####For Any Advice Please feel free to contact me at vet.m.mohamed@gmail.com

##Steps 

The following is the steps of running **EFA**

1- Testing the Assumptions

2- Specifying the number of factors 

3- Choosing the factor math and rotation type 

4- Exploring the Adequacy of EFA model 


####First of all lets load all libraries which may need to

```{r librares,message=FALSE,warning=FALSE}
library(psych) ; library(GPArotation)
library(bnstruct) ; library(knitr)
```

####- Now Lets import our data to work with 

```{r importing data }
efadf<-read.csv(file = "3 efa.csv")
head(efadf)
efadf<-efadf[,-21]
```

####here we have to take a look on the data structure and attr.for starting testing the assumptions

##1- testing the assumption


####it includes :

#####*Accurecy & Missing Value*

```{r str&attr}
str(efadf)

```

####as we see here we have a problem with the class of numeric value 

```{r class modification}
for(i in 1:(ncol(efadf))){
  efadf[,i]<-as.numeric(efadf[,i])
}
str(efadf)
```

####Great !!

####Now lets look at the summary of the data 

```{r data summary}
desc<-apply(efadf,2,function(x){
  (
    round(c(
    min=min(x,na.rm = T),
    max=max(x,na.rm = T),
    mean=mean(x,na.rm = T),
    na=sum(is.na(x))/nrow(efadf)
    ),2)
  )
})

print(desc)
```
####is there are any item has missing more than 5% ? 
```{r 5%}
any(desc["na",]>.05)
```
####great there is no item with more than 5% missing value 
####now test the missing for each individual
```{r miss ind}
ind.desc<-apply(efadf,1,function(x){
  round(c(sum(is.na(x))/ncol(efadf)),2)
})

any(ind.desc>.05)

which(ind.desc>.05)

ind.desc[which(ind.desc>.05)]
```
####woow it is so high percent 
####we will delete these members from the data 
```{r mem rem}
efadf<-efadf[-which(ind.desc>.05),]
```


####Now we have to impute the missing value 

```{r missing replace}
efadf<-knn.impute(as.matrix(efadf),k = 5)%>%data.frame()
```


####finally lets reverse the reversed items !!
```{r reverse}
key<-rep(1,ncol(efadf))

key[c(1,2,11,13,15,17,19,20)]<-key[c(1,2,11,13,15,17,19,20)]*-1

efadf<-reverse.code(keys = key,items = efadf)

```

####congrats: we finished this part **Acurecy and missing**

Now I am going to test some assumption related to factor analysis like:

1- **Additivity** 

2- **enough correlation between items**

3- **adequacy of sample **

```{r assump,warning=FALSE,message=FALSE}
#Additivity 

corr<-cor(efadf,method = "p")

apply(corr,2,function(x){
        any(x<1&x>=abs(.9))
})

#Ok !! there is no additivity 

#Now test the enough correlation using correlation bartelett test

cortest.bartlett(efadf)

#significant result indicate enough correlation 

#Finally KMO test for sample size adequacy 

KMO(efadf)

#here we can see that the data roughly adequate as it is >.8
```



Now it s the time to dive more in EFA 

##2- taking decision about the number of factors 

by following the next method 

1- **Parallel test** 

2- **Kaiser criterion**

3- **scree plot**

```{r fa number}
nf<-fa.parallel(x = efadf,fa = "fa")#suggestion of parallel test is 5 factors 

sum(nf$fa.values>.7) #Kaiser criterion suggest 3 factors

plot(nf) #scree plot suggest 2-3 factor 
```

####lets try 3 factors and look at the adequacy of the model 


```{r 3 fa}
efafit<-fa(efadf,nfactors = 3,rotate = "oblimin")
print(efafit)


#extracting the fit measure 

fitm<-c(RMS=efafit$rms,
RMSEA=efafit$RMSEA,
TLI=efafit$TLI,
CFI=(
        1-(efafit$STATISTIC-efafit$dof)/(efafit$null.chisq-efafit$null.dof)
))
print(fitm)
```

####the model here is poor but lets take a look on the loading and drawing the diagram

```{r}
fa.diagram(efafit)

efafit$loadings

```

####as we see here that all items are good regarding its loading to factors 

####So, lets try to repeat the analysis with 2 factors instead of 3 

```{r 2 factors}
efafit2<-fa(efadf,nfactors = 2,rotate = "oblimin")

print(efafit2)


#extracting the fit measure 

fitm2<-c(RMS=efafit2$rms,
RMSEA=efafit2$RMSEA,
TLI=efafit2$TLI,
CFI=(
        1-(efafit2$STATISTIC-efafit2$dof)/(efafit2$null.chisq-efafit2$null.dof)
))

print(fitm2)


fa.diagram(efafit2)

efafit2$loadings
```

####As we see here that there are a great distortion in the loading with a huge drop in CFI 

####so we will back to 3 factor model 


####The following codes for reporting in tables 

```{r tables}

fa.load<-matrix(efafit$loadings,ncol = 3,dimnames=list(c(paste0("q",1:20)),c("F1","F2","F3"))
)

fa.var<-matrix(efafit$Vaccounted,ncol = 3,dimnames = list(c("SS loading","Prop var","Cum var","Prop explain","Cum Prop"),c(paste0("MR",1:3))))

fa.cor<-matrix(efafit$score.cor,ncol = 3,dimnames = list(c(paste0("MR",1:3)),c(paste0("MR",1:3))))

efa.fit<-data.frame(fitm)

#Tables
kable(fa.var,format = "markdown",align = "c",digits = 3,caption = "Table1 : Variance Explanation for factors")
kable(fa.cor,format = "markdown",align = "c",digits = 3,caption = "Table2: the correlation between factors")
kable(fa.load,format = "markdown",align = "c",digits = 3,caption = "Table3: Item Factor Loading")

kable(efa.fit,format = "markdown",align = "c",digits = 3,caption = "Table4: Fit measures")

```

Finally lets Test the reliability of Each Factor by Alpha method

```{r alpha,warning=F,message=F}

#specifying the idx
f.idx<-list()
f.idx$f3<-c(5,13,6,19,9,10,11,17,20,4,8,7)
f.idx$f1<-c(1,12,14,18,16,2)
f.idx$f2<-c(15,3)
print(f.idx)


alpha(efadf[,f.idx$f1])$total #1st factor
alpha(efadf[,f.idx$f2])$total #2nd factor
alpha(efadf[,f.idx$f3])$total #3rd factor 

```

####Here we see that all alpha for all factors are above .80 , and it is a good indicator for the reliability of each factor 

##Congrats !!




