---
title: "ESEM Simulation"
author: "Mo'men Mohamed"
date: "March 7, 2019"
output: html_document
---

###loading libraries

```{r lib,message=F,warning=F}
library(lavaan)
library(bnstruct)
library(psych)
library(GPArotation)
```

###importing the scale

```{r scale imp}
df<-read.csv(file = "./data sets/c17.csv")

#looking at the class

apply(df,2,class)

#convering the class to numeric

for(i in 1:(ncol(df))){
        df[,i]<-as.numeric(df[,i])
}

str(df)

```

###Testing if there are any typo or missing 

```{r check}

#check by col
cch<-apply(df,2,function(x){
        c(min=min(x,na.rm = T),
          max=max(x,na.rm=T),
          na=sum(is.na(x))/nrow(df))
})

print(cch)

any(cch["na",]>.05)

#check by row
rch<-apply(df,1,function(x){
        c(na=sum(is.na(x)/ncol(df)))
})

any(rch>.05)


```
###Now there are some items need to be reversed 

```{r reverse}
key<-rep(1,ncol(df))
key[c(2,3,9,19,22,23)]<-key[c(2,3,9,19,22,23)]*-1

df<-reverse.code(keys =key,items = df )
df<-data.frame(df)
names(df)<-c(paste0("q",1:23))

```
###exploring some EFA assumption
####1- Additivity

####2- Correlation

####3- sample size adequacy

```{r assumption,warning=F,message=F}
#additivity

print(cor(df))

apply(df,2,function(x){
        any(x>=abs(.9)&x<1)
})
#Great!! there is no additivity 

#correlation enouph by bartlett test

cortest.bartlett(df)

#well the correlation test is significant , so the item is correlated enough 

#adequacy of sample size

KMO(df)

# the KMO score is >.9 and it indicates the adequacy of the sample size 

```

###Now lets test the number of factors

```{r No.Fa}
fa<-fa.parallel(df,fm = "ml",fa = "fa")
#parallel test suggesting 6 factor 

#lets see kaisar method eigen value>.7
sum(fa$fa.values>.7) # Kaisar suggestion of two 

#now see scree plot

plot(fa) #scree plot suggest 2 factor 
```

###So, we are going to choose the most parcimonious model --> 2 factor 

```{r fa}
efa<-fa(df,nfactors=2,rotate="geominQ",fm="ml",alpha=T)

#visualiz

fa.diagram(efa)
load<-zapsmall(round(matrix(efa$loadings,ncol = 2),2),3)
dimnames(load)<-list(names(df),c("f1","f2"))
print(load)

#removing Item 23
df<-df[,-23]

#running EFA Again
efa<-fa(df,nfactors=2,rotate="geominQ",fm="ml",alpha=T)
fa.diagram(efa)
load<-zapsmall(round(matrix(efa$loadings,ncol = 2),2),3)
dimnames(load)<-list(names(df),c("f1","f2"))
print(load)
```

###now lets run the normal CFA 

```{r CFA}
df<-data.frame(df)

cfamod<-"
f1=~q1+q2+q3+q4+q5+q6+q7+q9+q10+q12+q13+q14+q15+q16+q18+q19+q20+q21+q22
f2=~q8+q11+q17
"
fit1<-cfa(model = cfamod,data = df,estimator = "MLR")

summary(fit1,fit.measures=T)
```


###Now lets run ESEM Model

```{r ESEM}
esemmod<-vector()

for(i in 1:2){
        esemmod[i]<-paste0("f",i,"=~",paste0(c(load[,i]),"*",rownames(load),collapse = "+"))
}
print(esemmod)
esemmod<-paste0(esemmod,collapse = "\n")

esemfit<-cfa(esemmod,data = df,estimator="MLR")

summary(esemfit,fit.measures=T)
```

###the difference in CFI between the two model is **0.862** - **0.849** = **0.013** which is significant 