---
title: "ESEM and CFA Comparison in Measurement Invariance"
author: "Mo'men Mohamed"
date: "March 15, 2019"
output: html_document
---
This project is a Comparison between ESEM and CFA methods to test the measurement invariance of 213 Egyptian students on Big five personality scale (30 items) 

**Plz Feel Free to messege me for any advice or recommendations at vet.m.mohamed@gmail.com**
        
Loading the important package

```{r lib,message=F,warning=FALSE}
library(lavaan) ; library(GPArotation) ; library(knitr) ; library(knitLatex) ; library(readxl) ; library(ggplot2) ;library(tidyr) ; library(dplyr) ; library(magrittr) ; library(car) ; library(psych) ; library(kableExtra) ; library(semPlot) ; library(semTools)
```

Reading the data and importing it 

```{r import}
big<-read_excel(path = "./data sets/big5.xlsx")
```

Looking at the data structure 

```{r str}
str(big)
```

We need to remove "serial" and convert "sex" to factor Male and female


```{r modi}
big<-big[,-1]

big$Sex<-factor(x = big$Sex,levels = c(1,2),labels = c("male","female"))

kable(head(big),format = "markdown")
```


Now lets look at some summary for the data over all 

```{r summ}
summary(big) #Foor accurecy, out bound items and missing data 

```

Great , the there is no out of bound items or missing data , and the number of males and females is all most equal 


The normality and linearity assumption is critical for Factor analysis and SEM , So, Lets test it 


```{r Normality}
#Testing Multivariate Normality using QQplot From fake regression on random normal variable
set.seed(0124)
rn<-rnorm(n = nrow(big))
lm<-lm(rn~.,data = big[,-c(1,2)])

plot(lm,1) #Homogeniety

qqPlot(lm) # Normality and linearity

```

From The QQplot we can see that we can assume the normality and linearity of the data !!


The following steps include conducting EFA

Lets start :

Firstly lets test the assumptions : **Additivity**  ,  **sample adequacy**

```{r EFA assumptions}

#Additivity 

corr<-cor(x = big[,-c(1,2)],method = "p")

kable(corr,digits = 2,format = "markdown") #correlation table

corr%>%apply(MARGIN = 2,function(x){
any(x<1&x>=abs(.9))
}) #Testing if there are a correlation over .9 between the items to avoid multicolinearity 

#Enough correlation & Sample adequacy
cortest.bartlett(R = corr,n = nrow(big))$p.value#correlation bartelett test 

KMO(corr)$MSA #Kaiser-Meyer-Olkin test 

```

since the assumptions have been met ( sig cor Bartlett and MSA >.8 and no additive ), Now we can run EFA


AS there are a priori theory which assume the presence of 5 factors on which the items load in unique way , there is no need to test the number of factors 

but lets do it for fun !!

```{r No.FAC}
facno<-fa.parallel(x = big[,-c(1,2)],fm = "ml",fa = "fa") # as th data is normal , we choosed the Muximum likelihood method of Extraction 
```

The parallel test suggest 5 factors !! agreeing with the theory , but for fancy lets look  at scree plot & Kaiser Criterion 

```{r scree}
(facno$fa.values>.7)%>%sum() #Kaiser criterion suggest 5 Factors !!

plot(facno) #Scree plot Suggest 4 Factors with almost more one !!

```

As we see that Kaiser criterion of Eigen value suggest 5 factors and scree suggests 4 and almost one !! so we will consider those 5 !!

```{r factor Extraction}
fit<-fa(r = big[,-c(1,2)],nfactors = 5,rotate = "geominQ",fm = "ml") 
fa.diagram(fit,main = "diagram illustrating the EFA Model")
```

The theory says that the factors are uncorrelated and it is supposed to use *"Orthogonal"* rotation Neverthless we will use *oblique (oblimin geomin) Rotation* , which allows the factors to correlate and that to match the default rotation method in *M-Plus* for *ESEM* 
and also since the data is normal , so we used *maximum likelihood* Extraction method 

```{r loading & correltions}
fit$loadings%>%matrix(ncol = 5,dimnames = list(c(names(big[,-c(1,2)])),paste0("F",1:5)))%>%kable(format = "markdown",digits = 3,caption = "Factor Items loading",row.names = T)

fit$r.scores%>%matrix(nrow = 5,dimnames = list(c(paste0("f",1:5)),c(paste0("f",1:5))))%>%kable(format = "markdown",digits = 3,caption = "factor correlation")

```

at the tables above we see the factor-items loadings and factors correlations 

we need to know if there are any significant cross loading of any of the items 

```{r cross}

load<-fit$loadings%>%matrix(ncol = 5,dimnames = list(c(names(big[,-c(1,2)])),paste0("F",1:5)))%>%round(digits = 3)

load%>%apply(MARGIN = 2,FUN = function(x){
ifelse(test = x>=.3,yes = x,no = paste("-"))
})%>%matrix(ncol = 5,dimnames =list(c(names(big[,-c(1,2)])),paste0("F",1:5)))%>%kable(format = "markdown",digits = 3)

```

in the loading detection table we see that item O8 which is supposed to load on to factor 5 , loads on factor 1

here i feel free to delete this item especially as we have another 5 items in its factor

```{r item delete}
big<-big%>%select(-O8)
```

and then run EFA again 


```{r EFA again}
fit<-fa(r = big[,-c(1,2)],nfactors = 5,rotate = "geominQ",fm = "ml") 
fa.diagram(fit,main = "diagram illustrating the EFA Model")

load<-fit$loadings%>%matrix(ncol = 5,dimnames = list(c(names(big[,-c(1,2)])),paste0("F",1:5)))%>%zapsmall(digits = 2)%>%round(digits = 3)

load%>%apply(MARGIN = 2,FUN = function(x){
ifelse(test = x>=.3,yes = x,no = paste("-"))
})%>%matrix(ncol = 5,dimnames =list(c(names(big[,-c(1,2)])),paste0("F",1:5)))%>%kable(format = "markdown",digits = 3)

```

Now it is awesome , we have here a great factors structure 

So, lets test the reliability of each dimension using alpha

```{r alpha,message=F,warning=F}
F1<-big%>%select(E1,E2,E3,E4,E6,E7)
F2<-big%>%select(N1,N2,N3,N5,N8,N9)
F3<-big%>%select(A1,A2,A3,A4,A5,A6)
F4<-big%>%select(O2,O3,O4,O5,O6)
F5<-big%>%select(C3,C4,C5,C6,C7,C8)

dimenstions<-list(F1,F2,F3,F4,F5)

alphas<-data.frame(test=psych::alpha(check.keys = T,big%>%select(-c(Age,Sex)))$total[1],
                   F1=psych::alpha(F1)$total[1],
                   F2=psych::alpha(F2)$total[1],
                   F3=psych::alpha(F3)$total[1],
                   F4=psych::alpha(F4)$total[1],
                   F5=psych::alpha(F5)$total[1])

names(alphas)<-c("All test",paste("F",1:5))
alphas%>%kable(format = "markdown",digits = 2) # table contain row alpha for the all items and every factor
```

here the alpha values of each factor is roughly good , so we can claim here that the dimensions is reliable 


After finishing the EFA lets start in CFA and Confirming the structure of the model 

First of all we have to specify the model

```{r specify}
bigmodel<-"
f1=~E1+E2+E3+E4+E6+E7
f2=~N1+N2+N3+N5+N8+N9
f3=~A1+A2+A3+A4+A5+A6
f4=~O2+O3+O4+O5+O6
f5=~C3+C4+C5+C6+C7+C8
"
```


After specifying the model we will use `cfa()` function to run the model

```{r run the model}
bigfit<-cfa(model = bigmodel,data = big%>%select(-c(Age,Sex)))

summary(bigfit,standardized=T,fit.measures=T)

semPaths(bigfit,whatLabels = "std",layout = "tree")

```

The number of DF in our model is *406* and the number of parameters estimated are *73* so, the remaining df are *367* and we have an over identified model

The following is the table of all parameters estimated 

```{r para table}
parameterEstimates(bigfit,standardized = T)%>%kable(digits = 2,format = "markdown")

```


All Loaings are so good and all of them are significant but we have an issue with the correlation between the factors which supposed to be low , but here is so high and significant


Now lets look to the measurement fit indecies table

```{r fit}
c(fitmeasures(object = bigfit,fit.measures = c("chisq","df","cfi","tli","rmsea","rmsea.ci.upper","rmsea.ci.lower")))%>%round(digits = 3)%>%data.frame()%>%t()%>%kable(format = "markdown")

```

the model has a good measures as *CFI* is moderate as well as *TLI*

but *RMSEA* is very good with Excellent Interval 

The only issue in the model is the high correlation of the factors

**Now** Lets try building the model by **ESEM** Methods and see the difference between it and traditional **CFA**

```{r ESEM}
esemmodel<-vector()

for(i in 1:5){
esemmodel[i]<-paste0("f",i,"=~",paste0(c(load[,i]),"*",names(load[,1]),collapse = "+"))
}

esemmodel<-paste0(esemmodel,collapse = "\n") #Model Specification

esemfit<-cfa(model = esemmodel,data = big[,-c(1,2)]) #fitting the model

all.fit<-rbind(c(fitmeasures(object = bigfit,fit.measures = c("chisq","df","cfi","tli","rmsea","rmsea.ci.upper","rmsea.ci.lower")))%>%round(digits = 3)%>%data.frame()%>%t(),c(fitmeasures(object = esemfit,fit.measures = c("chisq","df","cfi","tli","rmsea","rmsea.ci.upper","rmsea.ci.lower")))%>%round(digits = 3)%>%data.frame()%>%t())%>%data.frame()

rownames(all.fit)<-c("CFA","ESEM")

all.fit%>%kable(format = "markdown")
```

SEE !! The fit measures improved a lot in case of ESEM method 

Now lets look at the correlation between the factors 

```{r factor corr}
data.frame(parameterestimates(object = esemfit,standardized = T))[180:189,c(1:3,11)]%>%kable(digits = 3,row.names = F,format = "markdown")
```

Here we can find the model match the theory as the factors are almost uncorrelated (Except f2~f3 , f2~f5 ) !!

**Now** I Think that it is time of exploring the measurement invariance using the two method 

##1- CFA Method 

```{r config inv,warning=F,message=F}

scalarfit<-cfa(model = bigmodel,data = big[,-c(1)],group = "Sex",group.equal = c("loadings","intercepts"))

measurementInvariance(model=bigmodel,data=big[,-c(1)],strict = T,group="Sex")

```

Here we see that we can't prove the scalar invariance because of the difference between it and metric invariance in **CFI** >.01 

So, we have to test the partial invariance using *freeing a parameter one-by-one* method 

```{r free test}

inter<-paste0(c(names(load[,1])),"~",1)

chi.diff<-list()

for(i in inter){
        test2<-cfa(model = bigmodel,
                   data = big[,-c(1)],
                   group = "Sex",
                   group.equal = c("loadings","intercepts"),
                   group.partial = i)
        chi.diff[[i]]<-c(i,fitmeasures(scalarfit,fit.measures = "chisq")-fitmeasures(test2,fit.measures = "chisq"))
}

chi.diff<-data.frame(chi.diff)%>%t()%>%data.frame()

names(chi.diff)<-c("intercept","chi-difference")

chi.diff<-chi.diff%>%mutate(`chi-difference`=as.character(`chi-difference`))%>%mutate(`chi-difference`=as.numeric(`chi-difference`))

chi.diff[order(chi.diff$`chi-difference`,decreasing = T),]%>%kable(digits = 3,format = "markdown",row.names = F,align = "c")
```

OK, the intercept of item `O3` has the big change in chisq , so, lets let it free and test partial measurement invariance 

```{r partial invariance,message=F,warning=F}
measurementInvariance(model=bigmodel,data=big[,-c(1)],strict = T,group="Sex",group.partial = "O3~1")
```

Congrats !! only `O3` is the differential item functioning , and the measurement is *partially invariant* with *Equal mean* across groups using TRADITIONAL CFA method 



##2- ESEM Method

lets judge on the measurement invariance using this new method and see how it acts differently!!
        
```{r esem invariance,warning=FALSE,message=FALSE}
measurementInvariance(model=esemmodel,data=big[,-1],strict = T,group="Sex")

```

See !! The measurement invariance is perfect at all levels with difference in **CFI** <.01 in all models and also with better fit measures. 

And here with ESEM we can stop at this point and say that there are a complete measure invariance in All Levels (Config. weak. strong. strict. means) of invariance measurement 

Finally lets make conclusion about our results of the two methods 

|Comparison facet | CFA | ESEM |
|:---------------|:----------------------------------------------|:----------------------------------------------|
|Model fit | The fit measures in CFA was poor compared to ESEM , CFI(.87)| The model here is very good with high CFI(0.974) and difference from CFA model is CFI > .01 and this is significant |
|Factors correlation | The factor correlation here doesn't match the theory as it is supposed to be uncorrelated | The correlation here is more better since most of factors uncorrelated |
|measurement invariance | The Model was good in proving the partial invarince (config , weak , strong , mean) | A conclusion of complete measurement invariance with better fit measures |





```{r}
fa<-factanal(x = big2,factors = 5,rotation = "geominQ",control = list(eps=.05))
fa$loadings%>%matrix(ncol = 5)%>%kable(format = "markdown")

```