```{r lib}
library(tidyverse)
library(lavaan)
library(knitr)
library(readxl)

```


loading the data

```{r load}

dat = lav_matrix_lower2full(c(.63, 
                                    .50, .60, 
                                    .48, .48, .58, 
                                    .47, .48, .51, .67))

means = c(5.17, 5.32, 5.40, 5.52)

names(means) = 
  rownames(dat) = 
  colnames(dat) = c("time1", "time2", "time3", "time4")

n=952
```

building the first model --> only intercept without variance 
```{r model1 }


model1<-"
#the model
i=~1*time1+1*time2+1*time3+1*time4

#zero Variance
i~~0*i

#fixing the residuals

time1~~r*time1
time2~~r*time2
time3~~r*time3
time4~~r*time4
"

fit1<-growth(model = model1,sample.cov = dat,sample.nobs = n,sample.mean = means)

summary(fit1,standardized=T,fit.measures=T,rsquare=T)

measures1<-fitmeasures(fit1,fit.measures = c("df","chisq","cfi","rmsea"))

measures1
```

in model 2 --> unconstrain the variance of intercept

```{r model 2}

model2<-"
#the model
i=~1*time1+1*time2+1*time3+1*time4

#fixing the residuals

time1~~r*time1
time2~~r*time2
time3~~r*time3
time4~~r*time4
"

fit2<-growth(model = model2,sample.cov  = dat,sample.nobs = n,sample.mean = means)

summary(fit2,standardized=T,fit.measures=T,rsquare=T)

measures2<-fitmeasures(fit2,fit.measures = c("df","chisq","cfi","rmsea"))

measures2
```

model 2 is better than model one as we permitted the sample to start from different points 

```{r model3}

model3<-"
#the model
i=~1*time1+1*time2+1*time3+1*time4
s=~0*time1+1*time2+2*time3+3*time4

#zero Variance
s~~0*s
i~~0*s

#fixing the residuals

time1~~r*time1
time2~~r*time2
time3~~r*time3
time4~~r*time4
"
fit3<-growth(model = model3,sample.cov  = dat,sample.nobs = n,sample.mean = means)

summary(fit3,standardized=T,fit.measures=T,rsquare=T)

measures3<-fitmeasures(fit3,fit.measures = c("df","chisq","cfi","rmsea"))

measures3
```

in model 3 we added slop but with zero variance and uncorrelated with intercept and it is better than the second model

```{r model4}
model4<-"
#the model
i=~1*time1+1*time2+1*time3+1*time4
s=~0*time1+1*time2+2*time3+3*time4

#fixing the residuals

time1~~r*time1
time2~~r*time2
time3~~r*time3
time4~~r*time4
"
fit4<-growth(model = model4,sample.cov  = dat,sample.nobs = n,sample.mean = means)

summary(fit4,standardized=T,fit.measures=T,rsquare=T)

measures4<-fitmeasures(fit4,fit.measures = c("df","chisq","cfi","rmsea"))

measures4

```

in model 4 we let the slop with intercept to vary 
and it is better than model 3

```{r model 5}

model5<-"
#the model
i=~1*time1+1*time2+1*time3+1*time4
s=~0*time1+1*time2+2*time3+3*time4

"
fit5<-growth(model = model5,sample.cov  = dat,sample.nobs = n,sample.mean = means)

summary(fit5,standardized=T,fit.measures=T,rsquare=T)

measures5<-fitmeasures(fit5,fit.measures = c("df","chisq","cfi","rmsea"))

measures5


```

in the model 5 it doesn't changed after model 4 and it means that the residuals are still constant 


```{r total fit}
total.fit<-rbind(measures1,measures2,measures3,measures4,measures5)

rownames(total.fit)<-c(paste0("model",1:5))

total.fit%>%kable("markdown",digits = 3)
```