---
title: "Mediation ( Bootstraping & lavaan SEM )"
author: "Mo'men Mohamed"
date: "April 10, 2019"
output: html_document
---


this is a tutorial about how to do basic mediation analysis 

For any recommendation of advices plz feel free to contact me : vet.m.mohamed@gmail.com

Need at first to load our libraries 

```{r,message=F,warning=F}
library(tidyverse)
library(mediation)
library(semPlot)
library(knitr)
library(caret)
library(lavaan)
```

Now lets import our data and inspect it 

```{r,eval=F}
download.file(url = 'http://static.lib.virginia.edu/statlab/materials/data/mediationData.csv',destfile = "./data sets/medanalysis.csv")
```

```{r}
data<-read.csv("./data sets/medanalysis.csv")

head(data)
```


here we have three variables x and y and the mediator M between x and y

first lets look at their features graphically befor starting analysis 

```{r}
plot(data)# seeing the corelation and linearity between each pairs


par(mfrow=c(1,3))
for(i in names(data)){
        density(data[,i])%>%plot(main=paste("density plot of variable",i,sep = " "))
} # seeing the normality of each variable

par(mfrow=c(1,3))
for(i in names(data)){
        d<-jitter(data[,paste(i)],factor = 10)
        qqnorm(d,main = paste("Normal QQ plot of",i,sep = " "))
} #seeing the normality and linearity using QQ plot 

```


Now lets dive directly to mediation 

1st model x~y

```{r}
fit1<-lm(Y~X,data=data)
summary(fit1)
```


the first model is significant 

now lets run the 2nd model m~x


```{r}
fit2<-lm(M~X,data=data)
summary(fit2)
```


great this is also significant 

Lets run the model 3 y~X+M --> adjusting for X


```{r}
fit3<-lm(Y~X+M,data)

summary(fit3)
```


here , M is significant amd X reduced to be completely non significant 

So we here in front of case of complete mediation 

but we still need to test if this mediation effect is significant or not 

So we will use bootstrabing method to test that 

```{r,cache=T}
medeffect<-mediate(model.m = fit2,model.y = fit3,sims = 1000,boot = T,boot.ci.type = "perc",treat = "X",mediator = "M")

summary(medeffect)
```


here we see ACME : Average causal mediation Effect is significant ,
and see that ADE : average direct effect is not significant 

This indicate that the mediation effect is significant !!


Now i want to test the nediation effect using Lavaan and structural equation modeling


```{r}
model<-"

Y~a*X+b*M
M~c*X
#indirect effect
ind:=b*c
#total effect
total:=a+(b*c)
#prop med
prop:=ind/total
"
```



now run the model through cfa() function


```{r}
medsem<-cfa(model = model,data = data,se="bootstrap")

summary(medsem,standardized=T,fit.measures=T,rsquare=T)

semPaths(object = medsem,whatLabels = "std")
```



we see here in the summary that the p value of X incase of presence of M is not significant and it is the same of model fit3 

also we see that the indirect Effect b*c is significant and the mediation effect is also significant 

Finaly we can see that the total effect is significant and it is so close to the total effect estimated in the bootstraping p.value in mediation effect model 




REGARDS