---
title: "MTMM"
author: "Mo'men Mohamed"
date: "March 29, 2019"
output: html_document
---

```{r lib,message=F,warning=F}
library(tidyverse)
library(lavaan)
library(knitr)
```

reading the data

```{r read}
dir<-file.path(choose.dir(),"mtmm example.csv")
dat<-read.csv(dir)

head(dat)%>%kable("markdown")
```

building the model
```{r model}
model1 = '

mlq =~ m1+m2+m3+m4+m5+m6+m8+m9+m10
pil =~ p3+p4+p8+p12+p17+p20

meaning =~ m1 + m2 + m5 + m10 + p4 + p12 + p17
purpose =~ m3 + m4 + m6 + m8 + m9 + p3 + p8 + p20

##fix the covariances
mlq~~0*meaning
pil~~0*meaning
mlq~~0*purpose
pil~~0*purpose
'

fit1<-cfa(model = model1,data=dat,std.lv=TRUE)

summary(fit1,standardized=T,fit.measures=T,rsquare=T)

measures1<-fitmeasures(fit1,fit.measures = c("df","chisq","cfi","rmsea"))

```


Building the second model 

```{r model2}

model2<-"
mlq =~ m1+m2+m3+m4+m5+m6+m8+m9+m10
pil =~ p3+p4+p8+p12+p17+p20
"


fit2<-cfa(model2,dat,std.lv=T)

summary(fit2,standardized=T,fit.measures=T,rsquare=T)

measures2<-fitmeasures(fit2,fit.measures = c("df","chisq","cfi","rmsea"))
```

Here the Delta of CFI between model 1 and model 2 is >,01 and it indicates convergent validity

Now lets see the third model at which we force the traits to be perfectly correlated 

```{r model 3}

model3<-"
mlq =~ m1+m2+m3+m4+m5+m6+m8+m9+m10
pil =~ p3+p4+p8+p12+p17+p20

meaning =~ m1 + m2 + m5 + m10 + p4 + p12 + p17
purpose =~ m3 + m4 + m6 + m8 + m9 + p3 + p8 + p20

##fix the covariances
mlq~~0*meaning
pil~~0*meaning
mlq~~0*purpose
pil~~0*purpose

##fix the traits to perfect correlation 

meaning~~1*purpose
"

fit3<-cfa(model3,dat,std.lv=T)

summary(fit3,standardized=T,fit.measures=T,rsquare=T)

measures3<-fitmeasures(fit3,fit.measures = c("df","chisq","cfi","rmsea"))

measures3
```

the model 3 is not worth that the first one and it indicate that we have a problem in discriminant validity of traits 

the fourth model force the methods to be uncorrelated 

```{r model4}

model4<-"
mlq =~ m1+m2+m3+m4+m5+m6+m8+m9+m10
pil =~ p3+p4+p8+p12+p17+p20

meaning =~ m1 + m2 + m5 + m10 + p4 + p12 + p17
purpose =~ m3 + m4 + m6 + m8 + m9 + p3 + p8 + p20

##fix the covariances
mlq~~0*meaning
pil~~0*meaning
mlq~~0*purpose
pil~~0*purpose

##fix the methods to perfect be uncorrelation 

mlq~~0*pil
"
fit4<-cfa(model4,dat,std.lv=T)

summary(fit4,standardized=T,fit.measures=T,rsquare=T)

measures4<-fitmeasures(fit4,fit.measures = c("df","chisq","cfi","rmsea"))

measures4
```

the model 4 didn't changed and it is as good as model 1 and it indicates Discriminant validity of Methods 

```{r fit collection}
fit.collection<-rbind(measures1,measures2,measures3,measures4)

rownames(fit.collection)<-c(paste0("measures",1:4))

fit.collection%>%kable("markdown",digits = 3)

```