---
title: "Pokemon data"
author: "Mo'men Mohamed"
date: "January 25, 2019"
output: html_document
---
```{r recommended libraries,message=FALSE}
library(readr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(factoextra)
library(NbClust)
library(gridExtra)
```


```{r importing data }
pok<-read_csv(file = "./data sets/Pokemon.csv")

head(pok)

```

looking at the data type for modification 
```{r data str}
str(pok)
```

Looking at the character variables if it is needed to convering to Factors instead

```{r chr to factor}
apply(pok[,c("Type 1","Type 2")],2,unique)
clust_var<-names(pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")])
```

I think that it will be useful to convert it to factors and the index to chr 

```{r convertion}

pok<-pok%>%mutate(`#`=as.character(`#`),Name=as.character(Name),`Type 1`=as.factor(`Type 1`),`Type 2`=as.factor(`Type 2`))

str(pok)
```

Now we will clustering the data 

and the first step to test the tendency of the data to be clustered 

and for that we will use a single function which test the *Hopkins statistics* and *VAT*

```{r clustering tendency,cache=TRUE}
scaled_pok<-scale(pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")])

tend<-get_clust_tendency(data = scaled_pok,n = 700)
tend
```

here we can see the Hopkins statistics is near to 1 **`r tend$hopkins_stat`** and it reveal that the data is clustered the correlation map indicate that it has inclination to 2 clusters

but we still need to determine it more accurtely as following 

```{r Kmean number of clusters, cache=TRUE}
#using NBclust package 

clus_num<-NbClust(pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")],method = "average",index = "gap")

plot(clus_num$All.index)

clus_num2<-NbClust(pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")],method = "average",index = "silhouette")

plot(clus_num2$All.index)

```

Now Lets try Kmeans Clusters 

```{r Kmeans,cache=TRUE}

set.seed(100)


kclust<-kmeans(x = pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")],centers = 2,nstart = 100)

pok$cluster<-as.factor(kclust$cluster)

plot(pok[,c("HP","Attack","Defense","Sp. Atk","Sp. Def","Speed")],col=pok$cluster)

```

lets see the distribution of types on the clusters

```{r}
table(pok$`Type 1`,pok$cluster)

table(pok$`Type 2`,pok$cluster)
```


Now lets explore all paramaters in light of clusters 


```{r HP }
tapply(pok$Attack,pok$cluster,function(x){
        return(c(mean=mean(x),
        median=median(x),
        sd=sd(x),
        IQR=IQR(x),
        CI=mean(x)+c(1,-1)*qnorm(.95)*sd(x)))
})

```