---
title: "Regression Coefficients SE and CI by Hand"
author: "Mo'men Mohamed"
date: "7/15/2020"
output: html_document
---
### This is a Tutorial to estimate SE and CI of Reg. Coef by hand
### To know how to estimate the OLS Coef by hand please visit [my Tutorial](https://rpubs.com/Momen/635174) , and it is a pleasure to know your feed back at [vet.m.mohamed@gmail.com](vet.m.mohamed@gmail.com)


 To Calculating standard error of b by hand we will use this formula : `var(b)=s^2(x'x)-1`


```{r}
library(tidyverse);library(readxl) 
```

importing and prepairing the data 

```{r}
dir<-"~/Statistical-practice/data sets/Ch3 application econometric analysis.xlsx"

dat<-read_xlsx(dir)

str(dat)

minus<-which(str_detect(dat$Ability,pattern = "−"))

dat$Ability<-str_remove_all(dat$Ability,"−")

dat$Ability<-as.numeric(dat$Ability)

dat$Ability[minus]<-dat$Ability[minus]*-1

names(dat)<-c("Person","Education","Wage","Experience","Ability","M_education","F_education","Siblings")

```


Look at it using regular function lm

```{r}
fit<-dat%>%lm(formula = Wage~Education+Ability)

summary(fit)
```

now get the (x'x)-1 matrix and S^2

```{r}


x<-cbind(C=rep(1,nrow(dat)),dat[,c("Education","Ability")])%>%as.matrix()

A<-solve(t(x)%*%(x))

S2<-(t(fit$residuals)%*%fit$residuals)/(nrow(dat)-length(coef(fit)))%>%as.vector()

varb<-S2[1,1]*diag(A) #The variance 

sdb<-sqrt(varb)%>%as.matrix() #standard error  

print(sdb)
```

Both Results seem to be identical 

Now calculate CI

```{r}
tk<-c(qt(p = c(.025,.975),df = (nrow(dat)-3)))%>%as.matrix()


cI<-coef(fit)+(sdb%*%t(tk))

print(cI)

confint(fit)

```

Again , Both are identical 

H testing 

```{r}
tvalue<-coef(fit)/sdb

print(tvalue)
#prop of t value

pt(q = tvalue,df = 12,lower.tail = F)
```


# Regards