---
title: "Sales Review"
author: "Mo'men Mohamed"
date: "12/31/2019"
output: powerpoint_presentation
---
# Sales Review For 2018-2019 
## including Compariaons between the Potential Accounts 

```{r lib,echo=FALSE,include=FALSE,cache=T}
library(knitr)
library(magrittr)
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r prep_data,echo=FALSE,include=F,cache=T}
sales<-read.csv(file = "/media/salsa/8F9E-E2B8/Sync/UBM/SALES ANALYSIS 018-019.csv")

head(sales)%>%kable()

for (i in 1:ncol(sales)) {
  class(sales[,i])%>%print()
}
#changing the classes
sales$Date<-as.Date(sales$Date)
sales$Qty<-as.numeric(sales$Qty)
sales$Unit.Price<-as.numeric(sales$Unit.Price)
sales$Net.Sales<-as.numeric(sales$Net.Sales)
sales$year<-as.factor(sales$year)
for (i in 1:ncol(sales)) {
  class(sales[,i])%>%print()
}
```



The First thing to pay attention to is to compare between the two years in **Total sales Value** as following:


```{r total,echo=F,cache=T}
sales%>%group_by(year)%>%summarize(Total_sales=sum(Net.Sales,na.rm = T))%>%ggplot(aes(x=year,y=Total_sales))+
  geom_col()+
  geom_label(aes(label=Total_sales))

```



As we See there are like *1 million and 800 thousands* difference between two years , so we need to break the sales down do analyse the gap and know where is the problem!

So, lets see each sector in the two years :


```{r prod,out.width='90%',echo=F,cache=T}
sales%>%group_by(year,Product.Type)%>%summarize(Total_sales=sum(Net.Sales,na.rm = T))%>%ggplot(aes(x=year,y=Total_sales,fill=Product.Type))+geom_col()+geom_text(aes(label=Total_sales,angle=45),check_overlap = T)+facet_grid(.~Product.Type)+
        theme(legend.position = "none")

```

As we see in the previous chart , the gap shrinked from 1800000 EGP to **900000 EGP**.

we need to know the the precise factor which dropped the sales like that , and also know the area of growth if it exists!

To know that lets First see the Accounts arranged in term of Total sales and see the relative importance of every one  in *2019* !

```{r,echo=F,cache=T}
sales%>%filter(year==2019)%>%group_by(Customer)%>%summarize(Sales=sum(Net.Sales,na.rm = T))%>%arrange(desc(Sales))%>%kable(format = "markdown")

```

**NOW** im going to throw the spot light on each singl Account of the most Potential one.

The aim is :

- To see the *drop* in sales **Product by Account**
- To see the *growth* also **Product by Account**

```{r prod cust,out.width="90%",echo=F,cache=T}

#Dina
sales%>%filter(Product.Type=="Add"&Customer=="Dina")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Dina :0")

#Copenhagen

sales%>%filter(Product.Type=="Add"&Customer=="Copenhagen")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Copenhagen")

#Delta Misr
sales%>%filter(Product.Type=="Add"&Customer=="Delta Misr")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Delta Misr")

#Enma’a
sales%>%filter(Product.Type=="Add"&Customer=="Enma'a")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm=T)*25)/1000)%>%ggplot(aes(x=year,y=Total))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Enma’a")

#Reta Keriasy

sales%>%filter(Product.Type=="Add"&Customer=="Reta Keriasy")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Reta Keriasy")

#Al Safa ( Al Maikana )

sales%>%filter(Product.Type=="Add"&Customer=="Al Safa ( Al Maikana )")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Al Safa ( Al Maikana )")

#Al- Pramos	

sales%>%filter(Product.Type=="Add"&Customer=="Al- Pramos")%>%group_by(year,Products)%>%summarize(Total=(sum(Qty,na.rm = T)*25)/1000)%>%ggplot(aes(x=year,y=Total,fill=Products))+geom_col()+geom_label(aes(label=Total))+facet_grid(.~Products)+ggtitle(label = "Al- Pramos")
```
