---
title: "The Power Of Matrix Algebra regarding OLS"
author: "Mo'men Mohamed"
date: "7/3/2020"
output: html_document
---
### It is a pleasure to know your feedback on [vet.m.mohamed@gmail.com]()


I Want to Show you the power of matrix algebra in estimating regression coefficients in R and coefficient of determination R2

**The data taken from *William H. Greene - Econometric Analysis, 7th Edition (2011, Prentice Hall)* , Ch3 application **

```{r,warning=FALSE,message=FALSE}
library(tidyverse);library(readxl);library(knitr)
```

**importing and prepairing the data **

```{r,warning=FALSE}
dir<-"~/Statistical-practice/data sets/Ch3 application econometric analysis.xlsx"

dat<-read_xlsx(dir)

str(dat)

minus<-which(str_detect(dat$Ability,pattern = "−"))

dat$Ability<-str_remove_all(dat$Ability,"−")

dat$Ability<-as.numeric(dat$Ability)

dat$Ability[minus]<-dat$Ability[minus]*-1

names(dat)<-c("Person","Education","Wage","Experience","Ability","M_education","F_education","Siblings")


dat%>%kable()
```


**splitting the data **

```{r}
X1<-cbind(C=rep(1,nrow(dat)),dat[,c("Education","Experience","Ability")])%>%as.matrix()
X2<-dat[,c("M_education","F_education","Siblings")]%>%as.matrix()
y<-dat$Wage%>%as.matrix()

X<-cbind(X1,X2)%>%as.matrix()

```


**regressing y on X1 and estimating coef. using OLS**

```{r}
coefX1<-solve(t(X1)%*%X1)%*%t(X1)%*%y

coefX1%>%kable(col.names = "coef")
```

**regressing y on X and estimating coef**

```{r}
coefX<-solve(t(X)%*%X)%*%t(X)%*%y
coefX%>%kable(col.names = "Coef")
```


**Regressing each of X2 on X1 and get the fitted values and Coef**

```{r}

#make identity matrix

I<-diag(1,nrow = nrow(dat),ncol = nrow(dat))

#make a projection matrix of X1

PX1<-(X1%*%solve(t(X1)%*%X1)%*%t(X1))

#Make a residual matrix of X1

MX1<-I-(X1%*%solve(t(X1)%*%X1)%*%t(X1))

#calculate Fitted Value of each Value of X2

X2starfitted<-PX1%*%X2

X2starfitted%>%kable()

#Calculate coef of each of X2 variables regressed on X1

coefX2star<-solve(t(X1)%*%X1)%*%t(X1)%*%X2
coefX2star%>%kable()

```

**Calculate R2 of y on both X1 and X2**

```{r}
#Make centering matrix
i<-rep(1,nrow(dat))
M0<-I-(i%*%solve(t(i)%*%i)%*%t(i))

#residual matrix Of total X
MX<-I-(X%*%solve(t(X)%*%X)%*%t(X))

#total residual of y on X1 and X2
e<-MX%*%y

#SSE calculation
s<-t(e)%*%e

#SST calculation
ym0y<-t(y)%*%M0%*%y

#R2 Calculation = 1-(SSE/SST)
R2<-1-(s/ym0y)

print(R2)
```

**computing R-square after removing the constant variable from X1**

```{r}
X_1<-X[,-1]

MX_1<-I-(X_1%*%solve(t(X_1)%*%X_1)%*%t(X_1)) #residual matrix

e_1<-MX_1%*%y #residual

s_1<-t(e_1)%*%e_1 #SSE

R2_1<-1-(s_1/ym0y) 

print(R2_1)
```

**computing R2 for constant term only **

```{r}
#the centering matrix is the same residual matrix made by constants only 

e_c<-M0%*%y #residual
s_c<-t(e_c)%*%e_c #SSE

R2_c<-1-(s_c/ym0y) #approximately zero

print(R2_c)
```

**regressing y on X1 and X2 after partialing effect of X1 on X2 (X2start) and calculating R2**

```{r}
X2star<-MX1%*%X2
Xstar<-cbind(X1,X2star)


Coef_star<-solve(t(Xstar)%*%Xstar)%*%t(Xstar)%*%y

Coef_star%>%kable(col.names = "Coef")

#residual matrix for star data 

Mstart<-I-((Xstar%*%solve(t(Xstar)%*%Xstar)%*%t(Xstar)))

estar<-Mstart%*%y

s_star<-t(estar)%*%estar

R2star<-1-(s_star/ym0y)

R2star
```

**fitting partitioned regression**

```{r}
#getting Coef of regressing y on X2 partialing out X1
coef_p<-solve(t(X2)%*%MX1%*%X2)%*%(t(X2)%*%MX1%*%y)

coef_p%>%kable(col.names = "coef")
#getting Coef of regressing y on X1 partialing out X2

coef_P2<-solve(t(X1)%*%X1)%*%t(X1)%*%(y-(X2%*%coef_p))

coef_P2%>%kable(col.names = "coef")
```


**Putting R2 together**

```{r}

#Table of R2

cbind(R2,R2_1,R2_c,R2star)%>%kable(col.names =c("R2","R2_1","R2_c","R2star") )

```

## **Good Luck**