---
title: "Therapist"
author: "Mo'men Mohamed"
date: "December 27, 2018"
output: html_document
---

Fisrstly need to load *rvest* library

```{r }
library(RSelenium)
```

other libraries may be useful

```{r}
library(magrittr)
library(stringr)
```


loading the remote driver

```{r}
rs<-rsDriver(browser = "firefox")
rmdriv<-rs$client
```




getting the main url of NY therapists 

```{r}
nymainurl<-"https://www.zencare.co/new-york-city/therapists"
```

navigating to the link
```{r}
rmdriv$navigate(nymainurl)
```

getting the name elements
```{r}
nameelement<-rmdriv$findElements(using = "xpath",value = "//h3[@class='therapistName']")

```

extracting the names

```{r}
names<-sapply(nameelement,function(x){
        x$getElementText()
})
```

extracting the name only 
```{r}
Names<-sapply(names,function(x){
        str_to_lower(x)%>%str_remove_all(pattern = "((\\,)(\\s)([:alpha:]{1,})){1,}")})
```

converting the list to data frame for easy manipulation 

```{r}
Names<-unlist(Names)
```


extracting href

```{r}
hrefelement<-rmdriv$findElements(using = 'xpath',value = "//div[@class='sqs-block-content']/div/a")
```

getting href

```{r}
href<-sapply(hrefelement,function(x){
        x$getElementAttribute("href")
})

href<-unlist(href)
```

making data frame of names and url

```{r}
Thirapistdata<-data.frame(Names,href)
```

extracting degree nodes

```{r}
degreeelement<-rmdriv$findElements(using = "xpath",value = "//span[@class='degree']")

```

getting the degree names

```{r}
Degree<-sapply(degreeelement,function(x){
        x$getElementText()
})

Degree<-unlist(Degree)
```

adding the dgree to data frame

```{r}
Thirapistdata<-cbind(Thirapistdata,Degree)

```


finding specialization element 

```{r}
specelement<-rmdriv$findElements(using="xpath",value="//p[@class='infoMobile']")

```

getting the specialization info 

```{r}
Spec<-sapply(specelement,function(x){
        x$getElementText()
})

Spec<-unlist(Spec)
```

adding it to the data frame 

```{r}
Thirapistdata<-cbind(Thirapistdata,Spec)
```